---
title: "P8105_HW6_YX2953"
author: "Vivian Xia"
date: "2024-12-01"
output: github_document
---

```{r}
# Load required libraries
library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(purrr)
library(broom)
library(knitr)
library(modelr)
library(rsample)
library(p8105.datasets)
```

Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
bootstrap_results <- weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    summary = map(models, broom::glance),
    results = map(models, broom::tidy)) %>%
  select(-models) %>%
  unnest(summary, results) %>% 
  select(id = .id, r.squared, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate) %>% 
  mutate(
    abs_prcp = abs(prcp),
    log_b1b2 = log(tmin * abs_prcp)
  ) %>% 
  select(id, r.squared, log_b1b2)
```

```{r}
bootstrap_results %>%
  ggplot(aes(x = r.squared)) + geom_density() + 
  labs(title = "Distribution of estimated r square")

LB_r = bootstrap_results %>% 
  pull(r.squared) %>%
  quantile(0.025)
UB_r =bootstrap_results %>%
  pull(r.squared) %>%
  quantile(0.975)

c(LB_r, UB_r)%>%
  knitr::kable(digits = 3)
```
The distribution shows a mild left-skew with its center approximately at 0.918. The lower and upper bounds of its 95% confidence interval, marked by the 2.5% and 97.5% quantiles, are `r LB_r |> round(3)` and `r UB_r |> round(3)`. This interval can be denoted as (`r LB_r |> round(3)`, `r UB_r |> round(3)`).


```{r}
bootstrap_results %>%
  filter(log_b1b2 != "NaN") %>%
  ggplot(aes(x = log_b1b2)) + geom_density() + labs(title = "Distribution of estimated log(beta1 * beta2)")

LB_b = bootstrap_results %>%
  filter(log_b1b2 != "NaN") %>%
  pull(log_b1b2) %>%
  quantile(0.025)

UB_b =bootstrap_results %>%
  filter(log_b1b2 != "NaN") %>% 
  pull(log_b1b2) %>%
  quantile(0.975)

c(LB_b, UB_b) %>%
  knitr::kable(digits = 3)
```

The distribution of is a left-skew, centering around -5.5. The distribution's 2.5% and 97.5% quantiles are `r LB_b |> round(3)` and `r UB_b |> round(3)`, respectively, defining a 95% confidence interval represented as (`r LB_b |> round(3)`, `r UB_b |> round(3)`).

Problem 2

```{r}
# Load and clean data
homicide_df <- read_csv(file = "data/homicide-data 2.csv", na = c("Unknown", "NA", "")) %>%
  mutate(
    reported_date = as.Date(as.character(reported_date), format = "%Y%m%d"),
    city_state = str_c(city, state, sep = ", "),
    solved = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>%
  
# Filter out specified cities and races
filter(
  !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
  victim_race %in% c("White", "Black")
  )
```


```{r}
# Logistic regression for Baltimore, MD
baltimore_df <- homicide_df %>%
  filter(city_state == "Baltimore, MD")

# Fit logistic regression model
logistic_model <- glm(solved ~ victim_age + victim_sex + victim_race, 
                      data = baltimore_df, family = binomial())

# Extract results for victim_sexMale and compute Odds Ratio
logistic_model_results <- logistic_model %>%
  broom::tidy(conf.int = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    OR = exp(estimate),
    CI_low = exp(conf.low),
    CI_high = exp(conf.high)
  ) %>%
  select(OR, CI_low, CI_high)

# Display results in a table
logistic_model_results %>% 
  kable(digits = 3, col.names = c("Odds Ratio (OR)", "95% CI Lower", "95% CI Upper"))
```
The estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is 0.426 and [0.324,0.558].

```{r}
# Logistic regression for all cities
city_models <- homicide_df %>%
  group_by(city_state) %>%
  nest() %>%
  # Apply logistic regression to each city
  mutate(
    model = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race, 
                            data = ., family = binomial())),
    results = map(model, ~ broom::tidy(.x, conf.int = TRUE) %>%
                    filter(term == "victim_sexMale") %>%
                    mutate(
                      OR = exp(estimate),
                      CI_low = exp(conf.low),
                      CI_high = exp(conf.high)
                    ) %>%
                    select(OR, CI_low, CI_high))
  ) %>%
  unnest(results) %>%
  select(city_state, OR, CI_low, CI_high)

# Display city-wise results in a table
city_models %>% 
  kable(digits = 3, col.names = c("City, State", "Odds Ratio (OR)", "95% CI Lower", "95% CI Upper"))
```


```{r}
# Step 4: Plot results for all cities
city_models %>%
  ggplot(aes(x = reorder(city_state, OR), y = OR)) +
  geom_point(size = 3, color = "blue") +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = 0.2, color = "darkgray") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", size = 1) +
  labs(
    title = "Odds Ratios for Solving Homicides by City",
    subtitle = "Comparison of Male vs Female Victims",
    x = "City, State",
    y = "Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )
```
The plot above displays the estimated odds ratios (ORs) and their confidence intervals for solving homicides, comparing male victims to female victims, across cities. Cities are organized by their ORs, showing the variation in how victim sex is associated with homicide resolution rates. An OR less than 1 indicates a lower likelihood of solving homicides for male victims compared to female victims and significant variation exists across cities, suggesting local factors may influence these outcomes.

Problem 3

```{r}
# Load the dataset
birthweight <- read_csv("data/birthweight.csv")

# Convert appropriate columns to factors
birthweight <- birthweight %>%
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform)
)
```


```{r}
# Fit the proposed regression model
model1 <- lm(bwt ~ blength + bhead + gaweeks + delwt + ppbmi + smoken + malform, data = birthweight)

# Summarize the model
summary(model1)

# Add residuals and predictions
birthweight <- birthweight %>%
  add_predictions(model1, var = "fitted") %>%
  add_residuals(model1, var = "residuals")

# Plot residuals vs. fitted values
ggplot(birthweight, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

```{r}
# Model 2: Length and gestational age as predictors
model2 <- lm(bwt ~ blength + gaweeks, data = birthweight)

# Model 3: Interactions between head circumference, length, and sex
model3 <- lm(bwt ~ bhead * blength * babysex, data = birthweight)
```

```{r}
# Cross-validation setup
set.seed(123)
cv_split <- vfold_cv(birthweight, v = 5)

# Function to compute RMSE for each model
compute_rmse <- function(model, split) {
  train_data <- analysis(split)
  test_data <- assessment(split)
  fit <- lm(model, data = train_data)
  predictions <- predict(fit, newdata = test_data)
  rmse <- sqrt(mean((test_data$bwt - predictions)^2))
  return(rmse)
}

# Compute RMSE for each model
rmse_model1 <- map_dbl(cv_split$splits, ~ compute_rmse(bwt ~ blength + bhead + gaweeks + delwt + ppbmi + smoken + malform, .x))
rmse_model2 <- map_dbl(cv_split$splits, ~ compute_rmse(bwt ~ blength + gaweeks, .x))
rmse_model3 <- map_dbl(cv_split$splits, ~ compute_rmse(bwt ~ bhead * blength * babysex, .x))

# Combine RMSE results into a single dataframe for plotting
rmse_results <- tibble(
  Model = rep(c("Model 1 (Proposed)", "Model 2 (Length + Gestational Age)", "Model 3 (Interactions)"), each = length(rmse_model1)),
  RMSE = c(rmse_model1, rmse_model2, rmse_model3)
)
```

```{r}
# Create a violin plot of RMSE for each model
ggplot(rmse_results, aes(x = Model, y = RMSE, fill = Model)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black") +
  labs(
    title = "RMSE Distribution Across Models",
    subtitle = "Comparison of Cross-Validated RMSE for Different Models",
    x = "Model",
    y = "RMSE"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )
```
The violin plot provides a detailed visualization of the distribution of RMSE (Root Mean Square Error) values across the different models evaluated during cross-validation. Model 1 includes multiple predictors like blength, bhead, gaweeks, etc. Model 2 focuses on blength and gaweeks.
Model 3 includes interactions between bhead, blength, and babysex. Model Performance:Model 1 likely has the lowest median RMSE, suggesting it performs better in predicting bwt compared to the others. Model 2 has a slightly higher median RMSE, indicating it is less accurate than Model 1 but likely simpler. Model 3 shows a wider spread of RMSE values, suggesting variability in its performance.
